Interoception Scale - Data Cleaning

### 2024-11-28 study1/analysis/1_cleaning.qmd
Updated cleaning script with attention checks

---
title: "Data Cleaning"
format: 
  html:
      code-fold: true
editor: visual
---

@@

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)

@@

#df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

df <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study1/data/rawdata_participants.csv")
```

@@

p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28) +
  # geom_line(aes(y = y), color = "orange", linewidth=2) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

@@

# did not add education disiplines

@@

## Ethnicity

# Ethnicity

p_eth <- df |>
  mutate(Ethnicity = fct_relevel(Ethnicity,"East Asian", "Mixed", "White", "Black", "persian", "South Asian", "Hispanic/Latino")) |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  #scale_fill_manual(values = c("East Asian"="#0571B0","Mixed"= "#d01c8b", "White"="#cc66cc", "Black"="#E66101", "persian" =  "#ca0020", "South Asian"="#5E3C99", "Hispanic/Latino"="#92C5DE")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6), angle = 45, hjust = 1),
    axis.title.x = element_blank())

@@

## Residence

@@

# Country participants are currently living

p_cou <- df |>
  mutate(Country = fct_relevel(Country,"United Kingdom", "Ireland", "USA", "Australia")) |>
  ggplot(aes(x = Country)) +
  geom_bar(aes(fill = Country)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Country of Residence", y = "Number of Participants", subtitle = "Participants per Country they are currently living") +
  theme_modern(axis.title.space = 15) +
  #scale_fill_manual(values = c("United Kingdom"="#ca0020","Ireland"= "#d01c8b", "United States"="#FDB863", "Australia"="#E66101"))+
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.7),  angle = 45, hjust =1),
    axis.title.x = element_blank())

p_cou

@@

p_country <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_country

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#E66101","Domains" = "#ca0020", "Random"="#d01c8b")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", fill = " ", x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  # geom_label(data = data.frame(x = mean(df$Experiment_Enjoyment) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Experiment_Enjoyment)))) +
  facet_wrap(~Condition) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    #axis.title.x = element_blank()
  )
p_feed

  #geom_line(aes(y = y), color = "orange", linewidth=2) +
#geom_vline(xintercept = mean(df$Experiment_Enjoyment), color = "#ca0020", linewidth=1.5) +
```

>>>>

---
title: "Interoception Scale - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
---

## Data Preparation

@@

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
```
```{r}
#| warning: false
#| code-fold: true
#| error: false

# Load the data

@@

df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")
```

## Attention Checks

df <- df |>
  dplyr::mutate(
    A1 = dplyr::case_when(
      Sexual_State_A == 6 | Sexual_State_A == 5 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A2 = dplyr::case_when(
      Anxious_SkinThermo_A == 0 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A3 = dplyr::case_when(
      Nociception_ColonBladder_A == 0 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A4 = dplyr::case_when(
      Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A5 = dplyr::case_when(
      Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A6 = dplyr::case_when(
      Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A7 = dplyr::case_when(
      Accuracy_Genital_A == 0 ~ "Pass",
      TRUE ~ "Fail"
    ),
    A8 = dplyr::case_when(
      Confusion_ColonBladder_A == 6 ~ "Pass",
      TRUE ~ "Fail"
    ),
  )

df$count.Fail <- apply(df[, c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "A8")], 1, function(x) length(which(x == "Fail")))

p_fail <- df |>
  ggplot(aes(x = count.Fail)) +
  geom_bar(aes(fill = count.Fail)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  # scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
   scale_fill_manual(values = c("0"="#0571B0","1"= "#d01c8b", "2"="#cc66cc", "3"="#E66101", "4" =  "#ca0020", "5"="#5E3C99", "6"="#92C5DE")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5)),
    axis.title.x = element_blank()
  )

p_fail

df <- df |> filter(count.Fail == 0)
```

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#E66101","Domains" = "#ca0020", "Random"="#d01c8b")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", fill = " ", x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  facet_wrap(~Condition) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
  )
p_feed

## Final Sample

::: {.panel-tabset}

@@

```{r p_demographics_age_gender}
p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),)
```

@@

```{r p_demographics_education}
# Did not add education disciplines

@@

## Ethinicity

```{r p_demographics_ethnicity}
p_eth <- df |>
  mutate(Ethnicity = fct_relevel(Ethnicity,"East Asian", "Mixed", "White", "Black", "persian", "South Asian", "Hispanic/Latino")) |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  scale_fill_manual(values = c("East Asian"="#0571B0","Mixed"= "#d01c8b", "White"="#cc66cc", "Black"="#E66101", "persian" =  "#ca0020", "South Asian"="#5E3C99", "Hispanic/Latino"="#92C5DE")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6), angle = 45, hjust = 1),
    axis.title.x = element_blank())

@@

## Country of Residence

```{r p_demographics_country}
p_cou <- df |>
 mutate(Country = fct_relevel(Country,"United Kingdom", "Ireland", "USA", "Australia")) |>
 ggplot(aes(x = Country)) +
 geom_bar(aes(fill = Country)) +
 scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
 scale_fill_viridis_d(guide = "none") +
 labs(title = "Country of Residence", y = "Number of Participants", subtitle = "Participants per Country they are currently living") +
 theme_modern(axis.title.space = 15) +
 scale_fill_manual(values = c("United Kingdom"="#ca0020","Ireland"= "#d01c8b", "United States"="#FDB863", "Australia"="#E66101"))+
 theme(
   plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
   plot.subtitle = element_text(size = rel(1.2), vjust = 7),
   axis.text.y = element_text(size = rel(1.1)),
   axis.text.x = element_text(size = rel(0.7)),
   axis.title.x = element_blank())

p_cou
```

```{r p_demographics_map}
p_map <- df |>
  mutate(Country = case_when(
    Country=="United States"~ "USA",
    Country=="United Kingdom" ~ "UK",
    TRUE ~ Country
  ))|>
  dplyr::select(region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  # mutate(n = replace_na(n, 0)) |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#cc66cc")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )

p_map
```

@@

# Create mapping from df's country names to world_map's country names. I think the map is not working because the names do not match.

country_mapping <- data.frame(
  df_name = c("United States", "United Kingdom"),
  world_map = c("USA", "UK") 
)

country_mapping

```
:::

```{r p_summary}
#| fig-width: 15
#| fig-height: 15

p_map / 
  (p_age + p_edu)
```

## Save

@@

write.csv(df, "../data/data_participants.csv", row.names = FALSE)
```
